Hierarchically Supervised Latent Dirichlet Allocation

### 2011-06-01 paper/model.tex
took a pass at inference.

HSLDA is a model for hierarchically, multiply-labeled, bag-of-words data.  We will refer to individual groups of bag-of-word data as documents.  Let $w_{n,d} \in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d = \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document $d$.  Let there be $D$ such documents and let the size of the vocabulary be $V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label $l \in \mathcal{L}$ has a parent $\mathrm{pa}(l) \in \mathcal{L}$ also in the set of labels.
 We will for exposition purposes assume that this label set has hard ``is-a'' parent-child constraints (explained later), although this assumption can be relaxed at the cost of more complicated inference.  Such a label hierarchy forms a multiply rooted tree.  Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.  Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label is applied to document $d$ or not.   In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.  In the applications we consider, only positive label applications are observed.

@@

Note that the choice of variables $a_{l,d}$ and how they are distributed were  driven at least in part by posterior inference efficiency considerations.  In particular, choosing probit-style auxiliary variable distributions for the $a_{l,d}$'s  yields conditional posterior distributions for both the auxiliary variables \eqref{eqn:a_l_d} and the regression coefficients \eqref{eqn:eta_l_cov}\eqref{eqn:eta_l_mean} which are analytic.  This simplifies posterior inference substantially.

>>>>

HSLDA is a model for hierarchically, multiply-labeled, bag-of-words data.  We will refer to individual groups of bag-of-word data as documents.  Let $w_{n,d} \in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d = \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document $d$.  Let there be $D$ such documents and let the size of the vocabulary be $V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label except root labels $l \in \mathcal{L}$ has a parent $\mathrm{pa}(l) \in \mathcal{L}$ also in the set of labels.
 We will for exposition purposes assume that this label set has hard ``is-a'' parent-child constraints (explained later), although this assumption can be relaxed at the cost of more complicated inference.  Such a label hierarchy forms a multiply rooted tree.  Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.  Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label is applied to document $d$ or not.   In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.  In the applications we consider, only positive label applications are observed.

@@

Note that the choice of variables $a_{l,d}$ and how they are distributed were  driven at least in part by posterior inference efficiency considerations.  In particular, choosing probit-style auxiliary variable distributions for the $a_{l,d}$'s  yields conditional posterior distributions for both the auxiliary variables \eqref{eqn:a_l_d} and the regression coefficients \eqref{eqn:regression_param_conditional} which are analytic.  This simplifies posterior inference substantially.
